\documentclass{diss}

\title{Practical Capacities, Know How, and Skill: The View from Reinforcement Learning}

\begin{document}
\maketitle

We know a lot. 
We know how to tie knots.
We know how to navigate across rooms, buildings, and cities.
We (or some of us) know how to make cheese.
We know, or can come to know, how to ride a bicycle or play the piano.

We have all sorts of know-how, and we have it in abundance.
And we are not alone in this: rats also know, in their own way, how to navigate across rooms, buildings, and cities and how to find cheese; with a little ingenuity, they might be taught to play some rudimentary piano.
Ants know how to find their way home after finding food, and bees know how to communicate the location of pollen to one another.

Clearly, these capacities are diverse.
But do they share a common nature?
Is there a common explanation of their possession or exercise?

According to \emph{intellectualism}, there is.
Intellectualists hold that knowing how to do something is knowing that something is true (namely, that such-and-such is a way to do that thing).

Intellectualism is the view that 

Intellectualists, however, would deny the rats, ants, and bees.
For according to intellectualists, to know how to do something is to know that something is true (namely, that such-and-such is a way to do that thing).
But rats, ants, and bees have no such knowledge.
According to intellectualism, they therefore lack know-how.
% Or, if they do, that knowledge is irrelevant to their possession of these practical capacities.
According to intellectualists, the rats do not, after all, know how to navigate rooms or find cheese (let alone play the piano).

\input{sharpening-the-question}

\input{computational_models.tex}

\input{representations.tex}
\end{document}
